Towards a Complete Model of the Negative Selection Process in Humans

### 2023-05-25 manuscript/manuscript.tex
new ms secs

\subsection*{Theory}

@@

\begin{align}
  \label{eq:}
  N_d &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
  R &= \frac{4N_d U s}{\exp(4 N_d s)-1}  & \text{\emph{ratchet equation}}
\end{align}

@@

\subsection*{Segment Simulations of the Diploid Model}

Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
haploid-model for a single region under equilibrium negative selection, we
first sought to confirm it could accurately predict the reduction in realistic
forward-time simulations of diploid populations with recombination. Using SLiM
\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
basepairs under varying levels of mutation and selection. We find a close
correspondence between the observed and predicted reduction in effective
population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
parameters including weak selection (Figure 1A), in contrast to classic BGS
theory. Furthermore, to investigate whether this accuracy was caused by the
model capturing the equilibrium fitness variance and ratchet rate, we also
measured these throughout the simulation. Again, we find their theory
accurately predicts both the ratchet rate (Figure 1B) and the genic fitness
variance (Figure 1C).

@@

Given the accuracy of the SC16 model in predicting the reduction in effective
population size and ratchet rate when equilibrium negative selection processes
were acting in single segment, we extended their model so that it could be
applied to whole-genome data. Our method, implemented in the software
\texttt{bgspy}, calculates the equilibrium additive genic fitness variance
$V_a$ and ratchet rate $R$ for each specified segment in the genome for a range
of mutation rates and selection coefficients and a fixed genome-wide
drift-effective population size $N$ (see Methods XXX). Genomic regions
putatively under negative selection (e.g. coding sequences or UTRs) and the
recombination rate map are specified as arguments so that the predicted values
accurately reflect local genomic details. Note that each equation is solved
independently, since here we assume no selective interference between segments.

To assess our genome-wide extension of the SC16 model, we simulated negative
selection for fixed mutation rates and selection coefficient on human
chromosome 10, using the HapMap recombination map (XXX) and in coding regions,
PhastCons regions, and UTRs as an approximate set of putatively conserved
segments. We then compared the reduction over an average of XXX simulation
replicates with the predicted reduction using our method.

We find that our theory closely matches simulations when selection is strong
(Figure 1XXX and Figure 1XXX). Since our theory is identical to classic BGS in
this domain, this provides the first genome-wide realistic forward-time
simulation confirmation of this theory. However, we find slight discrepancies
in regions with low recombination (Figure XXX). Second, we find our theory is
vastly more accurate than classic BGS theory when selection is weak ($2N_e s
\ll 1$; Figure 1XXX). However, when the strength of selection is at the
drift-barrier of $2N_e s \approx 1$), we find our theory has about a 10\%
relative error rate, which is slightly lower than classic BGS theory.

We first hypothesized that this discrepancy could be because slight selective
interference between segments

could be due to the build up of negative
linkage disequilibrium

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{figures/figure_1.pdf}
  \caption{Theory compared to simulation results. (A) The predicted reduction
      factor under classic B theory (dark gray line) and the diploid SC16 model
      (colored lines corresponding to mutation rate) compared to average
      reduction across XXX simulation replicates (points). (B) The predicted
      ratchet rate under the SC16 model scaled by mutation rate (colored lines)
      compared to the ratchet estimated from simulation (points). When $2Ns>1$,
      the ratchet rate is near zero. (C) The genic variance from simulations
      (points) against the predicted variance under the SC16 model (colored
      lines). As the ratchet begins to click, the genic variance is decreased
      from the level expected under strong BGS (dashed line). (D, bottom) The
      mean squared error (MSE) between whole-chromosome simulations and
      predicted classic B (dots), new B' (solid), and locally-rescaled B'
      (dashed) for different mutation rates (colors). The dashed horizontal
      line is the approximate theoretic minimum MSE  (D, top) The build up of
      negative linkage disequilibria around $2Ns=1$ in whole-chromosome
      simulations. (E) The average B across 100 chromosome 10 simulation
  replicates (gray) against different predictions, for parameters that
correspond to $2Ns < 1$, $2Ns = 1$, and $2Ns > 1$. }
  \label{fig:figure-1}
\end{figure}

@@

>>>>

\section*{Theory}

@@

\begin{align}
  \label{eq:main_eqns}
  N_d &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
  R &= \frac{4N_d U s}{\exp(4 N_d s)-1}  & \text{\emph{ratchet equation}}
\end{align}

@@

\subsection*{Simulations of a Segment under Negative Selection}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{figures/figure_1.pdf}
  \caption{Theory compared to simulation results. (A) The predicted reduction
      factor under classic B theory (dark gray line) and the diploid SC16 model
      (colored lines corresponding to mutation rate) compared to average
      reduction across XXX simulation replicates (points). (B) The predicted
      ratchet rate under the SC16 model scaled by mutation rate (colored lines)
      compared to the ratchet estimated from simulation (points). When $2Ns>1$,
      the ratchet rate is near zero. (C) The genic variance from simulations
      (points) against the predicted variance under the SC16 model (colored
      lines). As the ratchet begins to click, the genic variance is decreased
      from the level expected under strong BGS (dashed line). (D, bottom) The
      mean squared error (MSE) between whole-chromosome simulations and
      predicted classic B (dots), new B' (solid), and locally-rescaled B'
      (dashed) for different mutation rates (colors). The dashed horizontal
      line is the approximate theoretic minimum MSE. (D, top) The build up of
      negative linkage disequilibria around $2Ns=1$ in whole-chromosome
      simulations. (E) The average B map from 100 chromosome 10 simulation
  replicates (gray) against different predictions, for parameters that
correspond to $2Ns < 1$, $2Ns = 1$, and $2Ns > 1$. The chromosome shows 
the density of conserved sites and recombination map used in simulations. }
  \label{fig:figure-1}
\end{figure}

Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
haploid-model for a single region under equilibrium negative selection, we
first sought to confirm it could accurately predict the reduction in realistic
forward-time simulations of diploid populations with recombination. Using SLiM
\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
basepairs under varying levels of mutation and selection. We find a close
correspondence between the observed and predicted reduction in effective
population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
parameters including weak selection (Figure \ref{fig:figure-1}A), in contrast
to classic BGS theory. Furthermore, to investigate whether this accuracy was
caused by the model capturing the equilibrium fitness variance and ratchet
rate, we also measured these throughout the simulation. Again, we find the
diploid SC16 theory accurately predicts both the ratchet rate (Figure
1\ref{fig:figure-1}B) and the genic fitness variance (Figure
1\ref{fig:figure-1}C).

@@

Given the accuracy of the SC16 model in predicting the reduction in effective
population size and ratchet rate when equilibrium negative selection processes
were acting in single segment, we extended their model so that it could be
applied to whole-genome data. Our method, implemented in the software
\texttt{bgspy}, calculates the equilibrium additive genic fitness variance
$V_a$ and ratchet rate $R$ for each specified segment in the genome for a range
of mutation rates and selection coefficients and a fixed genome-wide
drift-effective population size $N$ (see Methods XXX). Genomic regions
putatively under negative selection (e.g. coding sequences or UTRs) and the
recombination rate map are specified as arguments so that the predicted values
accurately reflect local genomic details. Note that each equation is solved
independently, since here we assume no selective interference between segments.
We refer to the prediction reductions along the genome from our approach as B'
maps (analogous to B maps from \cite{McVicker2009-ax}).

To assess our B' maps, we simulated negative selection for fixed mutation rates
and selection coefficient on human chromosome 10 using realistic recombination
and conserved feature maps. Averaging over $n=100$ simulation replicates, we
estimated the simulation expected reduction map and compared this to the
predicted B and B' maps for the same parameters. We find that our B' maps and
the classic BGS theory B maps closely match simulations when selection is
strong (top row of Figure \ref{fig:figure-1}E). This provides the first
realistic forward-time simulation confirmation of classic background selection
theory. However, we find slight discrepancies in regions with low recombination
(Figure \ref{fig:figure-1}E). Second, we find our theory is vastly more
accurate than the classic B maps when selection is very weak ($2N_e s \ll 1$;
bottom row of Figure \ref{fig:figure-1}E), which is expected since BGS theory
does not work in this domain. Across all mutation and selection parameters
simulated, the relative error of the classic B maps is 14.6\% whereas the
relative error in the new B' maps is 5\%. Additionally, we find that the mean
squared error between simulations and B' maps is close to the theoretic lower
bound set by the coalescence variance \parencite[see Methods
XXX]{Tajima1983-gu}.

Nearly all of the error between the new B' maps and chromosome-wide simulations
occurs around the drift-barrier domain of $2Ns \approx 1$ (Figure
\ref{fig:figure-1}D). The higher error in this domain occurs across all
mutation rates. First, we hypothesized that the reason this occurred was the
when we numerically solve Equations \eqref{eq:main_eqns} for each segment, this
does does not consider the reduction due to selection at other linked segments.
In particular, we use $N=1000$ corresponding to the number of diploids in the
simulations rather than $B'(x) N$ at position $x$. To test this, we implemented
a locally-rescaled version of the B' maps, which numerically solves Equations
\eqref{eq:main_eqns} for each segment at position $x$ taking into account the
reduction experienced due to selection at other segments by setting $B'(x)N$ as
the population size (see Method XXX). We find the locally-rescaled B' maps
reduce the relative error from 5\% to 0.4\% and mean squared error (Figure
\ref{fig:figure-1}D, dashed colored lines), but does not entirely eliminate the
error in the $2Ns \approx 1$ domain.

Finally, we hypothesized that the remaining error is because of the build up of
negative linkage disequilibria between selected sites due to Hill--Robertson
interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. To assess
this possibility, we calculated the sum of all linkage disequilibria in our
chromosome-wide simulations. We find negative linkage disequilibria build ups
around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is greater
when mutation rates are higher (Supplementary Figure XXX). This is consistent
our view that the equilibrium fitness variance under the SC16 model is the
additive \emph{genic} fitness variance, which differs from the additive genetic
variance by the sum of linkage disequilibria between selected sites,
$\delta_{LD} = s^2 \sum_{i\nej} D_{i,k}$.

\subsection*{Estimates of Negative Selection in Humans}

@@

@@

\subsection*{Chromosome-wide Simulations}

, using the HapMap recombination map (XXX) and in coding regions, PhastCons
regions, and UTRs as an approximate set of putatively conserved segments

\subsection*{Local-rescaled }
We note that it is computationally intractable to solve this equation excluding
the contribution of the focal segment
